---
title: "dplyr Superhero"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: yes
    toc_float: yes
    keep_md: yes
---

## Learning Goals  
*At the end of this exercise, you will be able to:*    
1. Develop your dplyr superpowers so you can easily and confidently manipulate dataframes.  
2. Learn helpful new functions that are part of the `janitor` package.  

## Instructions
For the second part of lab today, we are going to spend time practicing the dplyr functions we have learned and add a few new ones. This lab doubles as your homework. Please complete the lab and push your final code to GitHub.  

## Load the libraries
```{r}
library("tidyverse")
library("janitor")
```

## Load the superhero data
These are data taken from comic books and assembled by fans. The include a good mix of categorical and continuous data.  Data taken from: https://www.kaggle.com/claudiodavi/superhero-set  

Check out the way I am loading these data. If I know there are NAs, I can take care of them at the beginning. But, we should do this very cautiously. At times it is better to keep the original columns and data intact.  
```{r}
superhero_info <- read_csv("data/heroes_information.csv", na = c("", "-99", "-"))
superhero_powers <- read_csv("data/super_hero_powers.csv", na = c("", "-99", "-"))
```

## Data tidy
1. Some of the names used in the `superhero_info` data are problematic so you should rename them here. Before you do anything, first have a look at the names of the variables. You can use `rename()` or `clean_names()`.    
```{r}
names(superhero_info)
```

```{r}
names(superhero_powers)
```

```{r}
superhero_info <- clean_names(superhero_info)
superhero_powers <- clean_names(superhero_powers)
```

## `tabyl`
The `janitor` package has many awesome functions that we will explore. Here is its version of `table` which not only produces counts but also percentages. Very handy! Let's use it to explore the proportion of good guys and bad guys in the `superhero_info` data.  
```{r}
tabyl(superhero_info, alignment)
```

1. Who are the publishers of the superheros? Show the proportion of superheros from each publisher. Which publisher has the highest number of superheros?  
```{r}
tabyl(superhero_info, publisher)
```
The publishers and proportions are listed above, the highest proportion appears to come form Marvel with about 53.9% of the comics.

2. Notice that we have some neutral superheros! Who are they? List their names below.  
```{r}
filter(superhero_info, superhero_info$alignment=="neutral") %>% select(name)
```
## `superhero_info`
3. Let's say we are only interested in the variables name, alignment, and "race". How would you isolate these variables from `superhero_info`?
```{r}
select(superhero_info, alignment, race)
```
Isolate them using select().
## Not Human
4. List all of the superheros that are not human.
```{r}
filter(superhero_info, race!="Human") %>% select(name) #any superhero that isn't fully human
```

## Good and Evil
5. Let's make two different data frames, one focused on the "good guys" and another focused on the "bad guys".
```{r}
good <- filter(superhero_info, alignment=="good")
good
```

```{r}
bad <- filter(superhero_info, alignment=="bad")
bad
```

6. For the good guys, use the `tabyl` function to summarize their "race".
```{r}
tabyl(good, race)
```

7. Among the good guys, Who are the Vampires?
```{r}
filter(good, race=="Vampire") %>% select(name)
bad
```
The only good vampires are Angel and Blade.
8. Among the bad guys, who are the male humans over 200 inches in height?
```{r}
filter(bad, height>200 & race=="Human" & gender=="Male")
```

9. Are there more good guys or bad guys with green hair?
```{r}
table(good$hair_color=="Green")
table(bad$hair_color=="Green")

```
More good guys have green hair than bad.

10. Let's explore who the really small superheros are. In the `superhero_info` data, which have a weight less than 50? Be sure to sort your results by weight lowest to highest.  
```{r}
small_supes <- filter(superhero_info, weight < 50) %>% arrange(weight)
small_supes
# arrange sorts by value desc does decreasing
```


11. Let's make a new variable that is the ratio of height to weight. Call this variable `height_weight_ratio`.    
```{r}
superhero_info %>% mutate(height_weight_ratio = height/weight) %>% select(name, height, weight, height_weight_ratio) %>% 
  arrange(desc(height_weight_ratio))
```

12. Who has the highest height to weight ratio?  
Groot has the highest height:weight ratio at a height of 701 and weight of 4.

## `superhero_powers`
Have a quick look at the `superhero_powers` data frame.
```{r}
glimpse(superhero_powers)
```


13. How many superheros have a combination of agility, stealth, super_strength, stamina?
```{r}
superhero_powers %>% filter(agility==TRUE, stealth==TRUE, super_strength==TRUE, stamina==TRUE) %>%  count
```
40 have a combination of all traits. 
## Your Favorite
14. Pick your favorite superhero and let's see their powers!  

```{r}
fav <- filter(superhero_info, superhero_info$name=="Dr Manhattan")

attributes <- filter(superhero_powers, hero_names=="Dr Manhattan") %>%  apply(1, function(data) names(which(data == T))) # had to look this up
attributes

```

```{r}
?filter
```

15. Can you find your hero in the superhero_info data? Show their info!  


```{r}
fav
```
Dr Manhattan is a Human/Cosmic character with no hair and blue skin from DC Comics.
## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.  