---
title: "Lab 6 Intro"
date: "`r Sys.Date()`"
output: 
  ioslides_presentation: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("jpeg")
```

## Setup
1. Go to iclicker.com and record your attendance.
2. Login to the lab computer (please don't use your personal computer).  
3. Navigate to github.com and login.  
4. Find the class repository "datascibiol" (https://github.com/jmledford3115/datascibiol).  
5. Download a copy the class repository to the desktop.  
6. Clone your repository to the desktop.  
7. Copy the files for today's lab from the class repository and paste them into **your** repository.  
8. Open today's lab in RStudio.  

## Review from last time
### *With a partner, discuss the following questions*
1. What are the characteristics of `tidy` data?  
2. What is the difference between `select` and `filter`?  
3. When is your first midterm?  

## Warm-up
1. Load the `bison.csv` data.
```{r}
bison <- read_csv("data/bison.csv")
```

2. What are the dimesions and structure of the data?
```{r}
summary(bison)
```
```{r}
glimpse(bison)
```


```{r}
dim(bison)
```
There are 8325 observations each with year of birth, rec year, rec month, rec day, animal code, animal sex, animal weight, and data code. 
3. We are only interested in code, sex, weight, year of birth. Restrict the data to these variables and store the dataframe as a new object.
```{r}
bison_summary <- select(bison, data_code, animal_sex, animal_weight, animal_yob)
```

4. Pull out the animals born between 1980-1990.

```{r}
bison_80s <- filter(bison_summary, between(animal_yob, 1980, 1990))
bison_80s
```

5. How many male and female bison are represented between 1980-1990?
```{r}
table(select(bison_80s, animal_sex))
```
There are 414 females are 21 males. 
6. Between 1980-1990, were males or females larger on average?
```{r}
#mean(select(filter(bison_80s, animal_sex=="F"),animal_weight ))
female_bison_weight_80s <- filter(bison_80s, animal_sex=="F") %>% select(animal_weight) 
female_bison_weight_80s
mean(female_bison_weight_80s$animal_weight, rm.na=T)
```

```{r}
male_bison_weight_80s <- filter(bison_80s, animal_sex=="M") %>% select(animal_weight) 
male_bison_weight_80s
mean(male_bison_weight_80s$animal_weight, rm.na=T)
```
On average, males are larger by 500 pounds. 